\documentclass[25pt, a0paper, portrait]{tikzposter}
\usetheme{Rays}
\usecolorstyle{Default}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb}

\definecolor{monFond}{RGB}{30, 40, 80}
\colorlet{backgroundcolor}{monFond}

\title{Makeitalive: Landscape picture animation}
\author{Mathurin Petit, Arthur Fournier}
\institute{CSC\_52002 - Generative AI Project}

\begin{document}
\maketitle

\begin{columns}
    \column{0.5}
    \block{1. Extraction Strategy}{
        We built a pipeline that automates the creation of image pairs $(I_A, I_B)$ from a youtube video called \textit{10 Hours Fantastic Views of Nature 4K with Relaxation Music}.
        The video is a compilation of drone views of landscapes which moves slowly forward. We extracted different pairs of images separated by a gap of 4 frames every 0.05 seconds to teach the model how to reconstruct successive images.
        We limited this number of pairs to 10000 for the first, small, dataset and 100 000 for the bigger one.\\
        We did not use the original format of the dataset. The video stream was reduced to 24 FPS and the extracted images were cropped to squares of 512 pixels and centered.
    }

    \block{2. Intelligent Data Filtering}{
        The Youtube video is a succession of landscape clip. To remove the pairs of pictures belonging to 2 different video clips, we used a distance to remove them from the dataset.
        We defined 2 different metrics :
        \begin{itemize}
            \item \textbf{Hist Distance}: Bhattacharyya ($>0.35$).
            \item \textbf{Thumbnail MAE}: Structural difference. Compares two lower images with their grey version ($>0.25$).
        \end{itemize}
        These 2 metrics are necessary because the MSE is no longer interesting as the images are not exactly the same.
        Combined toghether they ensure the model only learns real movement and not montage jumps.
    }

    \block{3. Architecture Parameters}{
        \begin{itemize}
            \item \textbf{Optimizer:} Adam ($10^{-4}$).
            \item \textbf{Loss:} Mean Squared Error (MSE) on flow.
            \item \textbf{Epochs:} 100 on a dedicated local dataset.
        \end{itemize}
    }

    \column{0.5}
    \block{4. Qualitative Analysis}{
        \begin{itemize}
            \item \textbf{Strengths:} Excellence in water and reflections.
            \item \textbf{Limitations:} Some artifacts on dense clouds due to flow ambiguity.
            \item \textbf{Edges:} \texttt{mode='reflect'} essential for aesthetics.
        \end{itemize}
    }

    \block{5. Performance (Loss Convergence)}{
        \begin{center}
        \begin{tabular}{|c|c|c|}
            \hline
            \textbf{Epoch} & \textbf{Training Loss} & \textbf{Validation Loss} \\
            \hline
            1 & 0.452 & 0.480 \\
            10 & 0.120 & 0.145 \\
            50 & 0.045 & 0.052 \\
            100 & 0.038 & 0.041 \\
            \hline
        \end{tabular}
        \end{center}
    }

    \block{6. Perspectives}{
        \begin{itemize}
            \item \textbf{Semantic Segmentation}: Apply specific flows to sky vs. water.
            \item \textbf{Multi-scale Inference}: Improve accuracy for details.
            \item \textbf{Temporal Refinement}: Auto-regressive prediction for smooth loops.
        \end{itemize}
    }

    \block{7. Impact}{
        The project demonstrates that a lightweight architecture can produce qualitative animations with very little training data.
    }
\end{columns}

\end{document}
